---
title: "Proposal Doc"
name: "Thomas, Kathleen, Dieter, Meg"
date: "10/10/2021"
output: html_document
---

## Introduction 
To visualize the effects of socioeconomic circumstance on American lives, we are exploring the relationship between life expectancy and high school graduation by state. A study from the Journal of Health and Social Behavior establishes that there is a clear correlation between a few socioeconomic factors and mortality, education is one factor that shows a particularly strong correlation (Phelan, Jo C., et al). We intend to explore and display this relationship by state, while considering the factor of racial background in some of our visuals. Our life expectancy data was collected by the Centers for Disease Control and Prevention, and our high school graduation rate data was collected by the National Center for Education Statistics. To narrow down our data, we will be exploring the data collected by these organizations from the year 2015.

The CDC collected this life expectancy data using a methodology designed for this particular project. There are three phases to this methodology, as described in "U.S. Small-area Life Expectancy Estimates Project: Methodology and Results Summary". Phase 1 entailed estimating death rates for census tracts with minimum pooled population sizes of 5,000 people, between the years of 2010 to 2015. Tracts with "acceptable" schedules of mortality were used to supplement tracts with age-specific death counts missing. Phase 2, the researchers used zero-truncated Poisson and negative binomial models and fit them to Phase 1 model tracts, resulting in model parameter estimates that they then used to predict age-specific death rates. Missing observed death rates due to lack of death counts were replaced with predicted values by combining demographic, socioeconomic, and geographic factors. In Phase 3, abridged life tables were calculated from the tracts. These tables include complete age-specific death counts (observed or predicted), along with age patters of mortality. These tables were then reviewed for reliability using a variety of statistical and demographic techniques. 

The National Center for Education and Statistics does not provide details regarding how they collected the data for this particular survey, but they do provide a manual containing their strict overarching guidelines that apply to each of their surveys. These standards aim to maintain the quality of the data being collected, the efficiency of sampling frames that are being reused in different projects, and ensure that response rates are high, and the privacy of survey participants is ensured at all times. The link to this guide can be found below:
https://nces.ed.gov/statprog/2012/pdf/Chapter3.pdf




## Data Set Overview
NOTE: The 4-year ACGR is the number of students who graduate in 4 years with a regular high school diploma divided by the number of students who form the adjusted cohort for the graduating class. From the beginning of 9th grade (or the earliest high school grade), students who are entering that grade for the first time form a cohort that is “adjusted” by adding any students who subsequently transfer into the cohort and subtracting any students who subsequently transfer out, emigrate to another country, or die. There are some differences in how states implemented the requirements for the ACGR, leading to the potential for differences across states in how the rates are calculated.  This is particularly applicable to the population of children with disabilities. Black includes African American, Hispanic includes Latino, Asian/Pacific Islander includes Native Hawaiian or Other Pacific Islander, and American Indian includes Alaska Native. Race categories exclude Hispanic origin unless specified. Variances of +/- 5% from year-to-year for any subgroup over 50 in cohort size were addressed with states through ED’s Coordinated Data Quality Review. With one exception, West Virginia, states provided explanations for the increase/decrease. The state comment was not accepted by the data steward (OESE) for California, Illinois, Louisiana, New Mexico, Texas, Utah and the District of Columbia. Those seven states and West Virginia received a letter from OESE to the state’s Chief State School Officer outlining data quality concerns that are expected to be addressed in future submissions.


Data for Asian, Hawaiian Native/Pacific Islander, and Two or more races students are being published by NCES for the first time in school year 2015–16.  Not all states report using all of the additional subgroups.  Subgroups that are not reported by a state appear as “– Not Available”. California and Utah report using the additional racial subgroup “FILIPINO”.  The cohort and resulting rate for the “Native Hawaiian/Pacific Islander” in California and Utah include students reported using the “FILIPINO” category.

```{r}
#install.packages("dplyr")
library(tidyverse)
```

```{r}
USA <- read_csv('Data/US_A.CSV')
NCES_DATA <- read_csv('Data/NCES Cleaned - 2015-16 ACGR.csv')
glimpse(USA)
glimpse(NCES_DATA)
```


## Works Cited

Arias, E, et al. U.S. Small-Area Life Expectancy Estimates Project :     Methodology and Results Summary. National Center for Health          Statistics, Sept. 2018.   https://www.cdc.gov/nchs/data/series/sr_02/sr02_181.pdf. 

Phelan, Jo C., et al. “‘Fundamental Causes’ of Social Inequalities in     Mortality: A Test of the Theory.” Journal of Health and Social       Behavior, vol. 45, no. 3, Sept. 2004, pp. 265–285,                   doi:10.1177/002214650404500303.





```{r}
paste('The dataset USA contains', NROW(USA), 'observations, and', NCOL(USA), 'variables. Our other dataset, NCES data contains',NROW(NCES_DATA), 'observations, and', NCOL(NCES_DATA), 'variables.')
```

## Data Analysis Plan
```{r}
state <- USA$STATE2KX
life_expect <- USA$`e(0)`
aggregate(life_expect,list(state), mean) # mean life expectancy per state
```

Within our data analysis, we will be using the high school graduation rate will be our explanatory variable and life expectancy will be our response variable. We will also be analyzing how data collected pertaining to different racial groups provides more insight into how populations are effected in different levels of severity. We will be using a variety of statistical methods to visualize this data including scatter plots with layered linear regressions, bar plots, and correlation matrices. This methods will helps us visualize the data and allow us to see any trends or other important information. We also hope to use the map functions within ggplot2 to provide geographic visualizations of the data we collect, which may result in conclusions about different regions of the United States. These statistical methods will ideally result in a clear correlation between the life expectancy and high school graduation rate in these states.

```{r}

```


